---
title: "df_apply()"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{df_apply()}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(RPackProj)

# Added so we can glimpse at the data
library(pillar)
```


## Overview

The df_apply() function takes a data frame and applies a given function to specified columns and returns a modified version of the original data frame with the function applied.


## Usage

``` r
df_apply(
  .data = ..., 
  .fun = ..., 
  .filter = ...,
  ...
)
```

## Arguments

#### .data

  This is a data frame that we want the function applied to
  
#### .fun

  This is the function we want applied to the given columns
  
#### .filter
  
  This is a function that can be applied to a column of a data frame. This filters down the columns we want the function applied to.

## Examples

```{r}
df1 <- data.frame(
  A = c(1, 2, 3),
  B = c(4, 5, 6),
  C = c("foo", "bar", "baz")
)

new_df1 <- df_apply(.data = df1, .fun = function(x) x * 2, .filter = is.numeric)

new_df1
```


```{r}
df2 <- data.frame(
  A = c(1.875, 2.654, 3.676),
  B = c(4.169, 5.456, 6.289),
  C = c(7.846, 8.186, 9.945)
)

new_df2 <- df_apply(.data = df2, .fun = round, .filter = is.numeric)

new_df2
```

```{r}
glimpse(population)

# Lets convert the population from millions into thousands
to_thousands <- function(x) {
  x * 1000
}

new_pop <- df_apply(.data = population, .fun = to_thousands, .filter = is.integer)

glimpse(new_pop)
```


```{r}
df3 <- data.frame(
   D = c(5, 5, 8, 3),
   E = c(4, 7, 4, 2),
   F = c(8, 10, 3, 4))
my_fun <- function(x, multiplier) x * multiplier
my_filter <- is.numeric
df_apply(.data = df3, .fun = my_fun, .filter = my_filter, multiplier = 2)
```


