---
title: "Plotly-Baseball"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Plotly-Baseball}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
install.packages("plotly")
library(RPackProj)
library(plotly)
library(tidyverse)
library(dplyr)
data(baseball)
```

```{r}
baseball <- baseball |>
  mutate(
    winPercentage = W/G,
    battingAvg = H/AB
  )
```


## Plotly Basics

Plotly is an R package that creates high quality, interactive plots and graphs. 

To get started, just call the plotting function `plot_ly()`. The most basic plot you can make in plotly is a scatterplot:

```{r message=FALSE, warning=FALSE}
tigers <- baseball |>
  filter(teamID %in% c('DET'))
plot_ly(data=tigers, x=~yearID, y=~W, type = 'scatter')
```

By including `marker()`, you can do other customization, like changing the point color.

```{r message=FALSE, warning=FALSE}
plot_ly(data=tigers, x=~yearID, y=~W, type = 'scatter', 
        marker = list(size = 7.5,
               color = 'orange')
        )
```

Use the pipe operator and you can include proper labeling. 

```{r message=FALSE, warning=FALSE}
plot_ly(data=tigers, x=~yearID, y=~W, type='scatter', 
        marker = list(size = 7.5,
               color = 'orange')
        ) |>
  layout(
    title = 'Detroit Wins since 1901',
    xaxis = list(title = 'Year'),
    yaxis = list(title = 'Wins')
  )
```

What if we wanted to 'connect the dots'? Just add `mode = 'lines'`

```{r}
plot_ly(data=tigers, x=~yearID, y=~W, type='scatter', mode = 'lines') |>
  layout(
    title = 'Detroit Wins since 1901',
    xaxis = list(title = 'Year'),
    yaxis = list(title = 'Wins')
  ) 
```

What if you want to have scatterplots of multiple teams?

```{r}
tigersCubsPhils <- baseball |>
  filter(yearID >= 1901) |>
  filter(franchID %in% c('DET', 'CHC', 'PHI')) 
plot_ly(data=tigersCubsPhils, x=~yearID, y=~W,
        type = 'scatter', color=~teamID)
```



### Animations 

One thing that makes plotly special is its addition of animations and interactivity.

In the example below, we'll look at the Detroit Tigers Winning Percentage over the years. 

```{r}
# Plotly's accumulate function. Have to run seperately because it is not 
#   built into the plotly package
accumulate_by <- function(dat, var) {
  var <- lazyeval::f_eval(var, dat)
  lvls <- plotly:::getLevels(var)
  dats <- lapply(seq_along(lvls), function(x) {
    cbind(dat[var %in% lvls[seq(1, x)], ], frame = lvls[[x]])
  })
  dplyr::bind_rows(dats)
}

# Create the figure and add the accumulate_by() function
fig <- tigers
fig <- fig %>% accumulate_by(~yearID)

# Start creating the figure
fig <- fig %>%
  plot_ly(
    x = ~yearID, 
    y = ~winPercentage,
    split = ~teamID,
    frame = ~frame, 
    type = 'scatter',
    mode = 'lines', 
    line = list(simplyfy = F)
  )

# Labels the x-axis
fig <- fig %>% layout(
  xaxis = list(
    title = "Year",
    zeroline = F
  ),
  # Labels the y-axis
  yaxis = list(
    title = "Winning Percentage",
    zeroline = F
  )
) 

# Sets animation settings
fig <- fig %>% animation_opts(
  frame = 500, 
  transition = 0, 
  redraw = FALSE
)

# Hide or Display the animation slider
fig <- fig %>% animation_slider(
  hide = F
)

# Set where the Play button will be on the plot
fig <- fig %>% animation_button(
  x = 1, xanchor = "right", y = 0, yanchor = "bottom"
)

# Display the plot
fig
```

Now we can see how the Detroit Tigers winning percentage has changed year over year since 1901. You can see they've had some good years (like 1934 and 1984), but also some really poor years (like 1952 and 2003). 


### 3-D Plots

Another cool feature that plotly has is its ability to create interactive 3-d plots.

```{r}
fig <- plot_ly(tigers, x = ~double, y = ~triple, z = ~HR, 
               marker=list(size = 3, color =~yearID, showscale=TRUE)) |>
  add_markers() 
fig
```

This plot isn't super helpful if you are doing an analysis of extra-base hits over time, but it does reveal some trends such as how triples have decreased over time while home runs have increased.

