---
title: "Visualizing Student Scores with Plotly"
author: "Richmond Yevudza"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Visualizing Student Scores with Plotly}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(RPackProj)
library(plotly)
```

## Load Package and Data

```{r message=FALSE, warning=FALSE}
data("student")
```

## Plotly Basics

Plotly is an R package that creates high-quality, interactive plots and graphs.

Create a Bar Chart using Plotly

To get started, just call the plotting function plot_ly(). The most basic plot you can make in Plotly is a bar chart: 

```{r message=FALSE, warning=FALSE}
plot_bar <- plot_ly(data = student, x = ~name, y = ~mark, type = "bar") %>%
  layout(title = "Student Scores",
         xaxis = list(title = "Student Names"),
         yaxis = list(title = "Scores"))

plot_bar
```

## Customizing Bar Chart Colors

By including marker(), you can do other customization, like changing the bar color.

```{r message=FALSE, warning=FALSE}
plot_bar_custom_colors <- plot_ly(data = student, x = ~name, y = ~mark, type = "bar",
        marker = list(color = 'orange')) %>%
  layout(title = "Student Scores",
         xaxis = list(title = "Student Names"),
         yaxis = list(title = "Scores"))

plot_bar_custom_colors
```

## Create a Scatterplot using Plotly

Scatterplots are another popular plot type. You can create a scatterplot by specifying type = 'scatter':

```{r message=FALSE, warning=FALSE}
plot_scatter <- plot_ly(data = student, x = ~name, y = ~mark, type = 'scatter', mode = 'markers') %>%
  layout(title = "Student Scores",
         xaxis = list(title = "Student Names"),
         yaxis = list(title = "Scores"))

plot_scatter
```

## Grouped Bar Chart by Gender

Grouped Bar Chart by Gender

```{r message=FALSE, warning=FALSE}
plot_grouped_bar <- plot_ly(data = student, x = ~name, y = ~mark, type = "bar", color = ~gender) %>%
  layout(title = "Student Scores by Gender",
         xaxis = list(title = "Student Names"),
         yaxis = list(title = "Scores"))

plot_grouped_bar
```

## Create a Box Plot using Plotly

Box plots are useful for visualizing the distribution of data. To create a box plot, specify type = 'box':

```{r message=FALSE, warning=FALSE}
plot_box <- plot_ly(data = student, y = ~mark, type = 'box') %>%
  layout(title = "Distribution of Student Scores",
         yaxis = list(title = "Scores"))

plot_box
```
